---
title: "Covid91 vaccine study Final2024"
author: "Wayne Jarvis"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

To ensure efficient data evaluation, it was essential to categorize the variables into four distinct groups. I created four variables labeled "M", "F", "gay", and "drug users" to separate them based on whether they should receive the Covid-19 vaccine. This allowed me to establish the effectiveness intervals. Each group was stored in its own variable for clarity and analysis.

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### Barplot

```{r}
dd2 <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot illustrates the relationship between Covid-19 infection rates, vaccine uptake, and placebo usage among the male population.The first graph displays the counts of individuals who either contracted Covid-19 or remained healthy, and whether they received the vaccine or a placebo. The second graph shows the percentage of people who had the placebo or the Covid-19 vaccine, and whether or not they contracted Covid-19. These groups are separated by their Covid-19 infection status and whether they received the placebo or vaccine.Among those who contracted Covid-19, the majority (around 75%) received the placebo, while a significantly smaller portion (about 25%) had the vaccine. Among the individuals who remained healthy, the distribution was more balanced, with a roughly 50/50 split between those who took the placebo and those who received the vaccine.The code used to generate this barplot starts with a code block labeled "dd2," which appears in the environment pane. This block contains four variables: treatment, infected, count, and percent. The second group, called "basicC," is also visible in the environment and contains 11 variables, which help create the first plot. The final code group, "basicCC," located in the files pane, contains 11 variables that are essential for generating the second, percentage-based plot.


## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions